Substitution rate responses to changes in effective population size.

### 2020-04-14 manuscript/supp-mat.tex
Maj manuscript, need more pedagogical writting

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{mdframed}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage[margin=60pt]{geometry}

@@

\begin{document}
\part*{Supplementary materials}
\tableofcontents
\section{$\bm{\omega}$ elasticity after a change in $\bm{\Ne}$}
\subsection{Genotype to phenotype map}
Define $\Nsite$ as the number of sites in the genotype sequence.
Each site can be in one of $\Nstate \geq 2$ states, where only $1$ state is defined the stable state, and $\Nstate - 1$ states are unstable.
For a given genotype sequence, define phenotype $0 \leq \x \leq 1$ as the current proportion of sites in the unstable state.
After a mutation, given that only one site can change at a time, the absolute change of $\x$ is either $0$ or $\dx=\sfrac{1}{\Nsite}$.
Define $\rho_{\x}(\dx)$ as the probability to get a change of phenotype equal to $\dx$, if the current phenotype is $\x$:\\
\begin{gather}
\begin{cases}
\dx &\text{ with probability } \rho_{\x}(\dx) = 1-x, \\
0 &\text{ with probability } \rho_{\x}(0) = \x \left[1 - \frac{1}{\Nstate - 1}\right], \\
-\dx &\text{ with probability } \rho_{\x}(-\dx) = \frac{\x}{\Nstate - 1}.\\
\end{cases} \label{eq:proba}
\end{gather}

\subsection{Selection coefficient}
$s(\x, \dx)$ is the selection coefficient of an effect $\dx$ if the current phenotype is $\x$:
\begin{align}
s(\x, \dx) &= \frac{f(\x + \dx) - f(\x)}{f( \x)}, \\
 &\simeq \frac{1}{f(\x)}\frac{ \partial f(\x) }{\partial \x} \dx, \\
 &\simeq \frac{ \partial \ln f(\x) }{\partial \x} \dx, \label{eq:s_from_fitness}
\end{align}
where $f( \x)$ is the Wrightian fitness of phenotype $x$. And thus we also have:
\begin{gather}
s(\x, -\dx) \simeq - s(\x, \dx) \text{ from eq. \ref{eq:s_from_fitness}} \label{eq:s_minus_deltax}, \\
\iff S(\x, -\dx) \simeq - S(\x, \dx), \label{eq:S_minus_deltax}
\end{gather}
where $S(\x\eq, \dx) = 4\Ne s (\x\eq, \dx)$ is the scaled selection coefficient.

\subsection{Probability of fixation}
The probability of fixation of a mutation with effect $\dx$, for a resident phenotype $\x$ is :
\begin{align}
p_{\text{fix}}(\x, \dx) &= \frac{ 2 s(\x, \dx)}{1 - \e^{-4\Ne s(\x, \dx)}}, \\
 &= \frac{ 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}}. \label{eq:pfix}
\end{align}
And in the case of neutral mutations, the probability of fixation is:
\begin{gather}
p_{\text{fix}}(\x, 0) = \frac{ 1}{2 \Ne} \label{eq:pfix0}.
\end{gather}
And the ratio of probability of fixation between selected and neutral mutations is:
\begin{align}
\frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} & = \frac{ 2 \Ne 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}},\\
& = \frac{S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \label{eq:pfixratio}.
\end{align}
\subsection{Equilibrium phenotype}
At equilibrium phenotype $\x\eq$, the expected selection coefficient of mutation that reached fixation must be $0$:
\begin{gather}
 0 = \E_{\dx} \left[ s(\x\eq, \dx) p_{\text{fix}}(\x\eq, \dx) \right], \\
\iff 0 = \frac{ 2 s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) + s(\x\eq, 0) \frac{ \rho_{\x\eq}(0)}{2\Ne} + \frac{ 2s(\x\eq, -\dx)^2}{1 - \e^{-S(\x\eq, -\dx)}} \rho_{\x\eq}(-\dx)\text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}}, \\
\Longrightarrow \frac{ 2s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) \simeq \frac{ - 2s(\x\eq, \dx)^2}{1 - \e^{S(\x\eq, \dx)}}   \rho_{\x\eq}(-\dx) \text{ from eq. \ref{eq:S_minus_deltax}}, \\
\iff \frac{ \rho_{\x\eq}(\dx)}{\rho_{\x\eq}(-\dx)} \simeq \e^{-S(\x\eq, \dx)} \frac{ \e^{-S(\x\eq, \dx)} - 1}{ \e^{-S(\x\eq, \dx)} \left( 1 - \e^{S(\x\eq, \dx)} \right)}, \\
\iff \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1) \simeq -S(\x\eq, \dx) \text{ from eq. \ref{eq:proba}} \label{eq:equilibrium}, \\
\iff \lambda_{K}(\x\eq) \simeq -S(\x\eq, \dx) \label{eq:equilibrium_lambda},
\end{gather}
where  $\lambda_{K}(\x\eq) = \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1)$.

@@

	\begin{tikzpicture}[
	declare function={
		omega(\x,\k)= \x*(2*(\x-1)*(ln(\k-1)+ln((1-\x)/\x))/(\k*(\x-1)+1)+ (\k-2)/(\k-1));
	},]
	\begin{axis}[
	width=0.65\textwidth,
	height=\axisdefaultheight,
	ylabel={$\dnds\eq$},
	xlabel={$\x\eq$},
	domain=0:1.0,
	ymin=0.0, ymax=1.1,
	samples=100,
	legend entries={$\Nstate=2$, $\Nstate=4$, $\Nstate=8$, $\Nstate=16$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.4)},anchor=north west}
	]
	\addplot[line width=2.0pt, color=BLUE]{omega(x, 2)};
	\addplot[line width=2.0pt, color=YELLOW]{omega(x, 4)};
	\addplot[line width=2.0pt, color=GREEN]{omega(x, 8)};
	\addplot[line width=2.0pt, color=RED]{omega(x, 16)};
	\end{axis}
	\end{tikzpicture}

@@

	\begin{tikzpicture}[
	declare function={
		domega(\x,\k)= 2*(\k*(\x-1)+1 + (\k*(\x-1)^2+2*\x-1)*(ln(\k-1) + ln((1-\x)/\x)))/(\k*(\x - 1) + 1)^2 + (\k-2) / (\k-1);
	},]
	\begin{axis}[
	width=0.65\textwidth,
	height=\axisdefaultheight,
	ylabel={$\frac{\der \dnds\eq}{\der \x\eq}$},
	xlabel={$\x\eq$},
	domain=0.0:1.0,
	ymin=0, ymax=6,
	samples=50,
	legend entries={$\Nstate=2$, $\Nstate=4$, $\Nstate=8$, $\Nstate=16$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.8)},anchor=north west}
	]
	\addplot[line width=2.0pt, color=BLUE]{domega(x, 2)};
	\addplot[line width=2.0pt, color=YELLOW]{domega(x, 4)};
	\addplot[line width=2.0pt, color=GREEN]{domega(x, 8)};
	\addplot[line width=2.0pt, color=RED]{domega(x, 16)};
	\end{axis}
	\end{tikzpicture}

@@

Moreover, with the approximation that $\left| 4 \Ne \frac{ \partial s(\x\eq, \dx) }{\partial \x\eq } \right| \gg \left| \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } \right|$, meaning that a change in phenotype causes a higher change in scaled selection coefficient than mutational bias, we have:
\begin{gather}
\frac{ \der \x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}},\\
\Longrightarrow \frac{ \der \x\eq}{\der \ln (\Ne)}  \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}. \label{eq:approx_dx_dlnNe}
\end{gather}
Together, these approximations leads to the following elasticity in equilibrium $\dnds$ after change in $\Ne$ as:
\begin{align}
\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}\label{eq:approx_dw_dlnNe}
\end{align}

@@

\subsection{Folded fraction}
All phenotype-fitness function considered below are log-concave, and as a result, $\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}$ is an decreasing function of $\x$; the less stable the protein already is, the stronger the purifying selection against additional destabilizing mutations. More precisely, fitness functions depends on the folded fraction of the protein of interest, which is given by the Fermi-Dirac distribution:
\begin{eqnarray}
P_{F} (\x) &=& \frac{1}{1 + e^{\beta(\alpha + \gamma \Nsite \x)}}
\end{eqnarray}
where $\x$ is the fraction of destabilizing mutations, each contributing a $\Delta \Delta G = \gamma$, and $\beta = 1 / kT$. Thus, $\alpha = \Delta G_0 < 0$ is the difference in free energy between folded and unfolded state when all sites are stable. $\Nsite \gamma$ is the expected change in $\Delta G$ when all sites are unstable. The misfolded fraction is  $P_{U} = 1-P_{F}$. In addition, $P_{F}$ is typically close to 1 (or $P_{U} \ll 1$), so that we can use a first-order approximation:
\begin{eqnarray}
P_{F}(\x) &=& 1 - P_{U}(\x) \\
&\simeq& 1 - e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
or equivalently
\begin{eqnarray}
P_{U}(\x) &\simeq& e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}

\subsection{Fitness equal to folded fraction}
A first model is to assume that the fitness is equal to the folded fraction \cite{Goldstein2013}:
\begin{gather}
f(\x) = \frac{1}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}}. \label{eq:fitness}
\end{gather}
The derivative of fitness w.r.t to phenotype is:
\begin{align}
\frac{\partial \ln (f(\x))}{\partial \x}  & = - \frac{\partial \ln \left( 1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)} \right)}{\partial \x} \text{ from eq. \ref{eq:fitness}}, \\
& = - \beta \Nsite \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \label{eq:df_dx},\\
& \simeq - \beta \Nsite \gamma \e^{\beta(\deltaGMin + \Nsite \gamma \x)}.
\end{align}
The equilibrium phenotype ($\x\eq$) is :
\begin{gather}
\lambda_{K}(\x\eq) = 4\Ne \beta \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x\eq)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x\eq)}}  \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.
\end{gather}

Using $\Ne=10^4$, $\beta=1.686$, $\deltaGMin = -118$, $\Nsite=300$, $\gamma=1$, we have the following :
\begin{center}
	\begin{tikzpicture}[
	declare function={
		lambda(\x,\k)= ln(\k-1) + ln((1-\x)/\x);
	}]
	\begin{axis}[
	width=0.65\textwidth,
	height=\axisdefaultheight,
	xlabel={$\x$},
	ymin=0.0, ymax=5.0,
	samples=100,
	legend entries={
		$4\Ne \beta \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}}$,
		$\lambda_{2}(\x)$,
		$\lambda_{4}(\x)$,
		$\lambda_{8}(\x)$,
		$\lambda_{16}(\x)$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.7,0.95)},anchor=north west}
	]
	\addplot[domain=0.0:0.38, line width=2.0pt, color=BLUE]{ 4 * 1000 * 1.686 *exp(1.686 * (-118 + 300 *x))};
	\addplot[domain=0.0:0.5,line width=2.0pt, color=YELLOW]{ lambda(x,2) };
	\addplot[domain=0.0:0.8,line width=2.0pt, color=GREEN]{ lambda(x,4) };
	\addplot[domain=0.0:0.9,line width=2.0pt, color=RED]{ lambda(x,8) };
	\addplot[domain=0.0:1.0,line width=2.0pt, color=LIGHTGREEN]{ lambda(x,16) };
	\end{axis}
	\end{tikzpicture}
\end{center}
Where in this example we can visually appreciate that the a change in phenotype causes a higher change in scaled selection coefficient than mutational bias (eq. \ref{eq:approx_dx_dlnNe}).
And the second derivative of fitness w.r.t to phenotype is:
\begin{align}
\frac{\partial^2 \ln (f(\x))}{\partial \x^2} & = - \beta \Nsite \gamma \frac{\partial}{\partial \x} \left( \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \right) \text{ from eq. \ref{eq:df_dx}}, \\
 & = - \beta \Nsite \gamma  \beta \Nsite \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{\left( 1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}\right)^2}, \\
 & = \frac{\beta \Nsite \gamma}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \frac{ \partial \ln f(\x) }{\partial \x} \text{ from eq. \ref{eq:df_dx}} \label{eq:df2_dx2}, \\
 & \simeq \beta \Nsite \gamma \frac{ \partial \ln f(\x) }{\partial \x} 
\end{align}
Finally, $\omega$ elasticity after a change in $\Ne$ is simply:
\begin{align}
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma} \text{from eq. \ref{eq:df2_dx2} and \ref{eq:approx_domega_dx}},
\end{align}
which is independent of $\x\eq$, meaning $\omega$ is linearly decreasing with $\Ne$ in log space.
This model, however, does not express the fact that selection is typically stronger for proteins characterized by higher levels of expression.

@@

A slight variation is to assume that the selective cost itself is proportional to the total amount of misfolded protein. For a given protein with expression level $y$:
\begin{eqnarray}
\ln f(\x) = - A y P_{U} (\x)
\end{eqnarray}
where $A$ is the cost per misfolded macromolecule. Then, 
\begin{eqnarray}
\frac{\partial \ln (f(\x))}{\partial \x} &\simeq& -A \beta \gamma \Nsite y  e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
or equivalently
\begin{eqnarray}
4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
Thus, under this model, an increase in the effective population size $\Ne$ or the expression level $y$ both have the same effect on the molecular evolutionary process undergone by the protein.
Moreover, $\omega$ elasticity after a change in $\Ne$ is the same as before:
\begin{align}
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.
\end{align}
\subsection{Translational errors}

An other variant account for translational errors. Translational errors occur at a rate $\rho$ per residue. These errors contribute additional destabilizing mutations, each with effect size $\delta x = 1/n$. The total number of translational errors per macromolecule is approximately Poisson distributed:
\begin{eqnarray}
\pi_k &=& e^{-\rho \Nsite} \frac{(\rho \Nsite)^k}{k!}
\end{eqnarray}
and the total selective cost is now an average over all possible values of $k$:
\begin{eqnarray}
\ln f(\x) &=& - A y \sum_k \pi_k e^{\beta(\alpha + \gamma\Nsite \x + \gamma k)}
\\
&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x)} \sum_k e^{-\rho n} \frac{(\rho n)^k}{k!} e^{\beta \gamma k}
\\
&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \Nsite (e^{\beta \gamma} -1 )}
\\
& \simeq & - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \beta \gamma \Nsite}
\\
& =& - A y  e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\end{eqnarray}
In words, the fitness function is the same the previous model, except that the trait $x$ (fraction of destabilizing mutations) is shifted by $\rho$, the mean fraction of additional mutations contributed by translation errors. This additional factor is independent of $x$, and as a result, the scaled selection strength is essentially the same, up to a proportionality constant (contributed by the shift):
\begin{eqnarray}
4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\\ &\propto& -4 \Ne y e^{\beta(\alpha + \gamma\Nsite \x)}
\end{eqnarray}
Moreover, $\omega$ elasticity after a change in $\Ne$ is again the same as before:
\begin{align}
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.
\end{align}

@@

To make the model tractable and compact, we assume that non-specific interactions are weak, i.e. $\epsilon_i << 1$ for all $i$. We then make a first-order approximation in the $\epsilon_i$'s. In addition, we use a mean-field approximation, such that, when considering a specific protein species $i$, we assume that all other proteins have the same fraction $\bar x$ of hydrophobic residues at their surface. The value of $\bar x$ could in principle be found using a self-consistent argument, essentially by (1) explicitly calculating the net substitution flux for protein $i$ with fraction $x_i$, under mean-field $\bar x$, and (2) expressing the constraint that this substitution process for protein $i$ is stationary at $x_i = \bar x$. This derivation is not conducted here, as it is not needed.

>>>>

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage[margin=60pt]{geometry}
\pdfinclusioncopyfonts=1
\captionsetup{width=0.85\textwidth}

@@

\begin{document}
\part*{Supplementary materials}
\tableofcontents
\section{$\bm{\omega}$ elasticity after a change in $\bm{\Ne}$}
\subsection{Genotype to phenotype map}
Define $\Nsite$ as the number of sites in the genotype sequence.
Each site can be in one of $\Nstate \geq 2$ states, where only $1$ state is defined the stable state, and $\Nstate - 1$ states are unstable.
For a given genotype sequence, define phenotype $0 \leq \x \leq 1$ as the current proportion of sites in the unstable state.
After a mutation, given that only one site can change at a time, the absolute change of $\x$ is either $0$ or $\dx=\sfrac{1}{\Nsite}$.
Define $\rho_{\x}(\dx)$ as the probability to get a change of phenotype equal to $\dx$, if the current phenotype is $\x$:\\
\begin{cases}
\dx &\text{ with probability } \rho_{\x}(\dx) = 1-x, \\
0 &\text{ with probability } \rho_{\x}(0) = \x \left[1 - \frac{1}{\Nstate - 1}\right], \\
-\dx &\text{ with probability } \rho_{\x}(-\dx) = \frac{\x}{\Nstate - 1}.\\
\end{cases} \label{eq:proba}
\subsection{Selection coefficient}
$s(\x, \dx)$ is the selection coefficient of an effect $\dx$ if the current phenotype is $\x$:
s(\x, \dx) &= \frac{f(\x + \dx) - f(\x)}{f( \x)}, \\
 &\simeq \frac{1}{f(\x)}\frac{ \partial f(\x) }{\partial \x} \dx, \\
 &\simeq \frac{ \partial \ln f(\x) }{\partial \x} \dx, \label{eq:s_from_fitness}
where $f( \x)$ is the Wrightian fitness of phenotype $x$. And thus we also have:
s(\x, -\dx) \simeq - s(\x, \dx) \text{ from eq. \ref{eq:s_from_fitness}} \label{eq:s_minus_deltax}, \\
\iff S(\x, -\dx) \simeq - S(\x, \dx), \label{eq:S_minus_deltax}
where $S(\x\eq, \dx) = 4\Ne s (\x\eq, \dx)$ is the scaled selection coefficient.
\subsection{Probability of fixation}
The probability of fixation of a mutation with effect $\dx$, for a resident phenotype $\x$ is :
p_{\text{fix}}(\x, \dx) &= \frac{ 2 s(\x, \dx)}{1 - \e^{-4\Ne s(\x, \dx)}}, \\
 &= \frac{ 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}}. \label{eq:pfix}
And in the case of neutral mutations, the probability of fixation is:
p_{\text{fix}}(\x, 0) = \frac{ 1}{2 \Ne} \label{eq:pfix0}.
And the ratio of probability of fixation between selected and neutral mutations is:
\frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} & = \frac{ 2 \Ne 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}},\\
& = \frac{S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \label{eq:pfixratio}.
\subsection{Equilibrium phenotype}
At equilibrium phenotype $\x\eq$, the expected selection coefficient of mutation that reached fixation must be $0$:
 0 = \E_{\dx} \left[ s(\x\eq, \dx) p_{\text{fix}}(\x\eq, \dx) \right], \\
\iff 0 = \frac{ 2 s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) + s(\x\eq, 0) \frac{ \rho_{\x\eq}(0)}{2\Ne} + \frac{ 2s(\x\eq, -\dx)^2}{1 - \e^{-S(\x\eq, -\dx)}} \rho_{\x\eq}(-\dx)\text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}}, \\
\Longrightarrow \frac{ 2s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) \simeq \frac{ - 2s(\x\eq, \dx)^2}{1 - \e^{S(\x\eq, \dx)}}   \rho_{\x\eq}(-\dx) \text{ from eq. \ref{eq:S_minus_deltax}}, \\
\iff \frac{ \rho_{\x\eq}(\dx)}{\rho_{\x\eq}(-\dx)} \simeq \e^{-S(\x\eq, \dx)} \frac{ \e^{-S(\x\eq, \dx)} - 1}{ \e^{-S(\x\eq, \dx)} \left( 1 - \e^{S(\x\eq, \dx)} \right)}, \\
\iff \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1) \simeq -S(\x\eq, \dx) \text{ from eq. \ref{eq:proba}} \label{eq:equilibrium}, \\
\iff \lambda_{K}(\x\eq) \simeq -S(\x\eq, \dx) \label{eq:equilibrium_lambda},
where  $\lambda_{K}(\x\eq) = \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1)$.

@@

 \includegraphics[width=0.8\textwidth, page=2] {figures.pdf}

@@

 \includegraphics[width=0.8\textwidth, page=3] {figures.pdf}

@@

Moreover, with the approximation that $\left| 4 \Ne \frac{ \partial s(\x\eq, \dx) }{\partial \x\eq } \right| \gg \left| \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } \right|$, meaning that a change in phenotype causes a higher change in scaled selection coefficient than mutational bias, we have:
\frac{ \der \x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne} \frac{\partial \lambda_K (\x\eq) }{\partial \x\eq } + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}},\\
\Longrightarrow \frac{ \der \x\eq}{\der \ln (\Ne)}  \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}. \label{eq:approx_dx_dlnNe}
Together, these approximations leads to the following elasticity in equilibrium $\dnds$ after change in $\Ne$ as:
\frac{ \der \omega\eq}{\der \ln (\Ne)} & \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}\label{eq:approx_dw_dlnNe}

@@

\subsection{Folded fraction}
All phenotype-fitness functions considered below are log-concave, and as a result, $\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}$ is an decreasing function of $\x$; the less stable the protein already is, the stronger the purifying selection against additional destabilizing mutations. More precisely, fitness functions depends on the folded fraction of the protein of interest, which is given by the Fermi-Dirac distribution:
\begin{eqnarray}
P_{F} (\x) &=& \frac{1}{1 + e^{\beta(\alpha + \gamma \Nsite \x)}}
\end{eqnarray}
where $\x$ is the fraction of destabilizing mutations, each contributing a $\Delta \Delta G = \gamma$, and $\beta = 1 / kT$. Thus, $\alpha = \Delta G_0 < 0$ is the difference in free energy between folded and unfolded state when all sites are stable. $\Nsite \gamma$ is the expected change in $\Delta G$ when all sites are unstable. The misfolded fraction is  $P_{U} = 1-P_{F}$. In addition, $P_{F}$ is typically close to 1 (or $P_{U} \ll 1$), so that we can use a first-order approximation:
\begin{eqnarray}
P_{F}(\x) &=& 1 - P_{U}(\x) \\
&\simeq& 1 - e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
or equivalently
\begin{eqnarray}
P_{U}(\x) &\simeq& e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
\subsection{Fitness equal to folded fraction}
A first model is to assume that the fitness is equal to the folded fraction \cite{Goldstein2013}:
f(\x) = \frac{1}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}}. \label{eq:fitness}
The derivative of fitness w.r.t to phenotype is:
\frac{\partial \ln (f(\x))}{\partial \x}  & = - \frac{\partial \ln \left( 1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)} \right)}{\partial \x} \text{ from eq. \ref{eq:fitness}}, \\
& = - \beta \Nsite \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \label{eq:df_dx},\\
& \simeq - \beta \Nsite \gamma \e^{\beta(\deltaGMin + \Nsite \gamma \x)}.
The equilibrium phenotype ($\x\eq$) is :
\lambda_{K}(\x\eq) = 4\Ne \beta \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x\eq)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x\eq)}}  \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.
Using $\Ne=10^4$, $\beta=1.686$, $\deltaGMin = -118$, $\Nsite=300$, $\gamma=1$, we have the following :
\includegraphics[width=0.8\textwidth, page=4] {figures.pdf}
Where in this example we can visually appreciate that the a change in phenotype causes a higher change in scaled selection coefficient than mutational bias (eq. \ref{eq:approx_dx_dlnNe}).
And the second derivative of fitness w.r.t to phenotype is:
\frac{\partial^2 \ln (f(\x))}{\partial \x^2} & = - \beta \Nsite \gamma \frac{\partial}{\partial \x} \left( \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \right) \text{ from eq. \ref{eq:df_dx}}, \\
 & = - \beta \Nsite \gamma  \beta \Nsite \gamma \frac{\e^{\beta(\deltaGMin + \Nsite \gamma \x)}}{\left( 1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}\right)^2}, \\
 & = \frac{\beta \Nsite \gamma}{1 + \e^{\beta(\deltaGMin + \Nsite \gamma \x)}} \frac{ \partial \ln f(\x) }{\partial \x} \text{ from eq. \ref{eq:df_dx}} \label{eq:df2_dx2}, \\
 & \simeq \beta \Nsite \gamma \frac{ \partial \ln f(\x) }{\partial \x} 
Finally, $\omega$ elasticity after a change in $\Ne$ is simply:
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma} \text{from eq. \ref{eq:df2_dx2} and \ref{eq:approx_domega_dx}},
which is independent of $\x\eq$, meaning $\omega$ is linearly decreasing with $\Ne$ in log space.
This model, however, does not express the fact that selection is typically stronger for proteins characterized by higher levels of expression. 

@@

A slight variation is to assume that the selective cost itself is proportional to the total amount of misfolded protein. For a given protein with expression level $y$:
\begin{eqnarray}
\ln f(\x) = - A y P_{U} (\x)
\end{eqnarray}
where $A$ is the cost per misfolded macromolecule. Then, 
\begin{eqnarray}
\frac{\partial \ln (f(\x))}{\partial \x} &\simeq& -A \beta \gamma \Nsite y  e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
or equivalently
\begin{eqnarray}
4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite \x)}
\end{eqnarray}
Thus, under this model, an increase in the effective population size $\Ne$ or the expression level $y$ both have the same effect on the molecular evolutionary process undergone by the protein.
Moreover, $\omega$ elasticity after a change in $\Ne$ is the same as before:
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.
\subsection{Translational errors}
An other variant account for translational errors. Translational errors occur at a rate $\rho$ per residue. These errors contribute additional destabilizing mutations, each with effect size $\delta x = 1/n$. The total number of translational errors per macromolecule is approximately Poisson distributed:
\begin{eqnarray}
\pi_k &=& e^{-\rho \Nsite} \frac{(\rho \Nsite)^k}{k!}
\end{eqnarray}
and the total selective cost is now an average over all possible values of $k$:
\begin{eqnarray}
\ln f(\x) &=& - A y \sum_k \pi_k e^{\beta(\alpha + \gamma\Nsite \x + \gamma k)}
\\
&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x)} \sum_k e^{-\rho n} \frac{(\rho n)^k}{k!} e^{\beta \gamma k}
\\
&=& - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \Nsite (e^{\beta \gamma} -1 )}
\\
& \simeq & - A y  e^{\beta(\alpha + \gamma\Nsite \x) + \rho \beta \gamma \Nsite}
\\
& =& - A y  e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\end{eqnarray}
In words, the fitness function is the same the previous model, except that the trait $x$ (fraction of destabilizing mutations) is shifted by $\rho$, the mean fraction of additional mutations contributed by translation errors. This additional factor is independent of $x$, and as a result, the scaled selection strength is essentially the same, up to a proportionality constant (contributed by the shift):
\begin{eqnarray}
4 \Ne \frac{\partial \ln (f(\x))}{\partial \x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite (x + \rho))}
\\ &\propto& -4 \Ne y e^{\beta(\alpha + \gamma\Nsite \x)}
\end{eqnarray}
Moreover, $\omega$ elasticity after a change in $\Ne$ is again the same as before:
\frac{ \der \dnds\eq}{\der \ln (\Ne)} & \simeq - \frac{1}{\beta \Nsite \gamma}.

@@

To make the model tractable and compact, we assume that non-specific interactions are weak, i.e. $\epsilon_i << 1$ for all $i$. We then make a first-order approximation in the $\epsilon_i$'s. In addition, we use a mean-field approximation, such that, when considering a specific protein species $i$, we assume that all other proteins have the same fraction $\bar x$ of hydrophobic residues at their surface. The value of $\bar x$ could in principle be found using a self-consistent argument, essentially by (1) explicitly calculating the net substitution flux for protein $i$ with fraction $x_i$, under mean field $\bar x$, and (2) expressing the constraint that this substitution process for protein $i$ is stationary at $x_i = \bar x$. This derivation is not conducted here, as it is not needed.
Using these approximations, we can re-express the conservation of total mass as:
\begin{eqnarray}
C &=& [i] + (m-1) [i] \frac{C}{C_0} \, e^{\beta (\alpha + \gamma \Nsite (\bar x + x_i))}
\end{eqnarray}
